Loop-Aware Optimizations in PyPy's Tracing JIT

### 2012-08-16 talk/dls2012/paper.tex
explain the "type analysis"

\begin{itemize}
    \item \lstinline{new} creates a new object.
    \item \lstinline{get} reads an attribute of an object.
    \item \lstinline{set} writes to an attribute of an object.
    \item \lstinline{guard_class} is a precise type check. It typically precedes
    an (inlined) method call and is followed by the trace of the called method.
\end{itemize}

@@

\revc{
In general, the paper is over-long on generalities and too short on details.
For example, the description of the basic technique at the beginning of section
5 is the third time the idea is explained at basically the same level of detail
(the others are in section 2 and section 4).  In contrast, the optimizations
applied rely on a simple type analysis, but this is only briefly alluded to.
}

@@

\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
 & CPython & Psyco & PyPy  & PyPy & GCC \\
 &         &       & no LP &      & -O3 \\
\hline
conv3(1e5) & 77.89 & 9.52 & 1.77  & 0.68 &  0.59 \\
\hline
conv3(1e6) & 77.15 & 9.58 & 1.69  & 0.77 &  0.74 \\
\hline
conv3x3(1000) & 233.54 & 125.40 & 0.57 & 0.27 & 0.25 \\
\hline
conv3x3(3) & 234.45 & 126.28 & 0.60 & 0.31 & 0.28 \\
\hline
conv5(1e5) & 122.54 & 16.67 & 1.86  & 1.05 &  0.65\\
\hline
conv5(1e6) & 125.77 & 16.80 & 1.92 & 1.09 &  0.80 \\
\hline
dilate3x3(1000) & 232.51 & 125.85 & 3.89 & 3.69 & 0.25 \\
\hline
sobel(1000) & 181.49 & 95.05 & 0.71 & 0.42 & 0.20 \\
\hline
sqrt(Fix16) & 744.35 & 421.65 & 3.93  & 2.14  & 0.96 \\
\hline
sqrt(float) & 24.21 & 5.52 & 1.36 & 1.00 & 0.98\\
\hline
sqrt(int) & 20.84 & 1.78 & 2.26 & 1.82  & 0.80 \\
\hline
\hline
Variations & - & - & $\pm 0.03$ & $\pm 0.01$ & $\pm 0.01$ \\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrices of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrices, only the number of rows are
  specified.} 
\end{figure}

@@

Benchmarks were run on Intel i7 M620 @2.67GHz with 4M cache and 8G of RAM
using Ubuntu Linux 11.4 in 32bit mode.
The machine was otherwise unoccupied. We use the following software
for benchmarks:

\begin{itemize}
\item PyPy 1.5
\item CPython 2.7.2
\item Psyco 1.6 with CPython 2.6.6
\item GCC 4.4.5 shipped with Ubuntu 11.4
\end{itemize}

>>>>

\begin{itemize}
    \item \lstinline{new} creates a new object.
    \item \lstinline{get} reads an attribute of an object.
    \item \lstinline{set} writes to an attribute of an object.
    \item \lstinline{guard_class} is a precise type check. It typically precedes
    an (inlined) method call and is followed by the trace of the called method.
    The type that the guard checks for is the one that the variable had during
    tracing.
\end{itemize}

@@

\revc{
In general, the paper is over-long on generalities and too short on details.
For example, the description of the basic technique at the beginning of section
5 is the third time the idea is explained at basically the same level of detail
(the others are in section 2 and section 4).
}

@@

\begin{figure}
\begin{center}
{\smaller
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
 & CPython & PyPy  & PyPy & GCC & luajit & luajit \\
 &        & no LP &       & -O3 &        & no LP \\
FFT(1024, 32768) & 469.07 & 20.83 +- 0.02 & 12.73 +- 0.01 & - & 2.76 +- 0.01 & 4.42 +- 0.01\\
\hline
FFT(1024,32768) & - & - & - & 1.40 +- 0.04 & - & -\\
\hline
FFT(1048576, 2) & 58.93 & 4.12 +- 0.01 & 2.05 +- 0.00 & - & 1.08 +- 0.01 & 1.33 +- 0.01\\
\hline
FFT(1048576,2) & - & - & - & 0.83 +- 0.02 & - & -\\
\hline
LU(100, 4096) & 1974.14 & 32.22 +- 0.14 & 13.39 +- 0.03 & - & 1.52 +- 0.01 & 8.57 +- 0.01\\
\hline
LU(100,4096) & - & - & - & 1.33 +- 0.04 & - & -\\
\hline
LU(1000, 2) & 955.31 & 14.98 +- 0.22 & 5.99 +- 0.21 & - & 0.67 +- 0.01 & 3.99 +- 0.01\\
\hline
LU(1000,2) & - & - & - & 0.65 +- 0.04 & - & -\\
\hline
MonteCarlo(268435456) & 618.89 & 20.60 +- 0.05 & 15.33 +- 0.08 & 1.69 +- 0.05 & 2.82 +- 0.00 & 3.92 +- 0.01\\
\hline
SOR(100, 32768) & 1458.12 & 8.24 +- 0.00 & 2.66 +- 0.00 & - & 1.31 +- 0.01 & 2.02 +- 0.00\\
\hline
SOR(100,32768) & - & - & - & 1.76 +- 0.04 & - & -\\
\hline
SOR(1000, 256) & 1210.45 & 6.48 +- 0.00 & 2.10 +- 0.00 & - & 1.08 +- 0.01 & 1.63 +- 0.00\\
\hline
SOR(1000,256) & - & - & - & 1.49 +- 0.02 & - & -\\
\hline
SparseMatMult(1000, 5000, 262144) & 371.66 & 24.25 +- 0.04 & 16.52 +- 0.04 & - & - & -\\
\hline
SparseMatMult(1000,5000,262144) & - & - & - & 1.84 +- 0.03 & 4.53 +- 0.02 & 9.64 +- 0.02\\
\hline
SparseMatMult(100000, 1000000, 1024) & 236.93 & 17.01 +- 0.01 & 8.75 +- 0.08 & - & - & -\\
\hline
SparseMatMult(100000,1000000,1024) & - & - & - & 1.20 +- 0.03 & 2.42 +- 0.01 & 7.19 +- 0.01\\
\hline
conv3(100, nil) & - & - & - & - & 0.18 +- 0.00 & 0.70 +- 0.00\\
\hline
conv3(1000, nil) & - & - & - & - & 0.12 +- 0.01 & 0.67 +- 0.01\\
\hline
conv3(1e5) & - & - & - & 0.52 +- 0.04 & - & -\\
\hline
conv3(1e6) & - & - & - & 0.60 +- 0.03 & - & -\\
\hline
conv3(array(1e5)) & 50.14 & 1.09 +- 0.01 & 0.49 +- 0.01 & - & - & -\\
\hline
conv3(array(1e6)) & 49.20 & 1.13 +- 0.02 & 0.51 +- 0.00 & - & - & -\\
\hline
conv3x3(1000) & - & - & - & 0.17 +- 0.04 & - & -\\
\hline
conv3x3(1000, 1000) & - & - & - & - & 0.09 +- 0.01 & 0.49 +- 0.01\\
\hline
conv3x3(1000000, 3) & - & - & - & - & 0.13 +- 0.00 & 0.53 +- 0.00\\
\hline
conv3x3(3) & - & - & - & 0.19 +- 0.03 & - & -\\
\hline
conv3x3(Array2D(1000000x3)) & 139.81 & 0.70 +- 0.00 & 0.21 +- 0.00 & - & - & -\\
\hline
conv3x3(Array2D(1000x1000)) & 138.95 & 0.70 +- 0.00 & 0.20 +- 0.00 & - & - & -\\
\hline
conv5(100, nil) & - & - & - & - & 0.21 +- 0.01 & 0.87 +- 0.01\\
\hline
conv5(1000, nil) & - & - & - & - & 0.17 +- 0.01 & 0.84 +- 0.00\\
\hline
conv5(1e5) & - & - & - & 0.55 +- 0.02 & - & -\\
\hline
conv5(1e6) & - & - & - & 0.58 +- 0.03 & - & -\\
\hline
conv5(array(1e5)) & 74.65 & 1.22 +- 0.00 & 0.64 +- 0.00 & - & - & -\\
\hline
conv5(array(1e6)) & 77.94 & 1.26 +- 0.00 & 0.68 +- 0.01 & - & - & -\\
\hline
dilate3x3(1000) & - & - & - & 0.17 +- 0.03 & - & -\\
\hline
dilate3x3(1000, 1000) & - & - & - & - & 0.09 +- 0.00 & 0.48 +- 0.01\\
\hline
dilate3x3(Array2D(1000x1000)) & 137.52 & 4.35 +- 0.01 & 3.91 +- 0.02 & - & - & -\\
\hline
sobel(Array2D(1000x1000)) & 104.02 & 0.49 +- 0.00 & 0.21 +- 0.00 & 0.17 +- 0.03 & - & -\\
\hline
sobel\_magnitude(1000, 1000) & - & - & - & - & 0.24 +- 0.01 & 0.60 +- 0.01\\
\hline
sqrt(Fix16) & 463.46 & 5.12 +- 0.00 & 2.96 +- 0.00 & 1.34 +- 0.03 & 1.14 +- 0.00 & 12.80 +- 0.04\\
\hline
sqrt(float) & 14.99 & 1.37 +- 0.00 & 0.89 +- 0.00 & 0.85 +- 0.04 & 0.83 +- 0.01 & 1.06 +- 0.01\\
\hline
sqrt(int) & 13.91 & 3.22 +- 0.02 & 2.65 +- 0.00 & 1.25 +- 0.03 & 0.83 +- 0.01 & 1.06 +- 0.00\\
\hline
\end{tabular}
}
\end{center}
\label{fig:benchmarks}
\caption{Benchmark Results in Seconds. Arrays of length $10^5$ and
  $10^6$ and matrices of size $1000\times 1000$ and $1000000 \times
  3$ are used. The one used in each benchmark is indicated in
  the leftmost column. For the matrices, only the number of rows are
  specified.} 
\end{figure}

@@

Benchmarks were run on Intel Xeon X5680 @3.33GHz with 12M cache and 16G of RAM
using Ubuntu Linux 11.4 in 64bit mode.
The machine was otherwise unoccupied. We use the following software
for benchmarks:

\begin{itemize}
\item PyPy 1.9
\item CPython 2.7.1
\item GCC 4.5.2 shipped with Ubuntu 11.4
\item LuaJIT 2.0 beta, commit ID 0dd175d9e711f039c663d35e96c149b705bcf450
\end{itemize}
